---
title: "single allele shared"
author: "Laura Zhou"
date: "7/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Load required library
  
```{r warning=FALSE, echo = TRUE, results = 'hide', warning = FALSE, message = FALSE}
library(data.table)
library(plyr)
library(ggplot2)
library(ggpubr)
theme_set(theme_bw())
library(ggpointdensity)
library(ggrepel)
library(stringr)
```

## Training data (post duplication removal)

Note: All five splits of SA combined is the full set of SA data. Training data is the first 4 splits


```{r}
# train = fread("../../../data/SA_data/by_split/train_v4_el_single_HLA_0.txt.gz")
train = fread("../../data/SA_data/train_v4_el_single_HLA.txt.gz")
head(train)
dim(train)
length(unique(train$HLA))
table(train$HLA)
table(train$binder, train$HLA)

train$pepseq = str_replace_all(train$peptide,"X","")
train$pep_length = nchar(train$pepseq)
head(train)

table(train$pep_length)
table(train$binder, train$pep_length)
```

## Looking at the HLA identified as having large proportion binders below 0.1

```{r}
train1 = train[which(train$HLA == "HLA-A25:01" |train$HLA == "HLA-A68:01"
                     | train$HLA == "HLA-C03:03"),]

table(train1$HLA, train1$pep_length, train1$binder)
table(train1$HLA)
```

## Training data: Overlapping of peptides in HLA

```{r}
train.data0 = fread("../../data/NetMHCpan4_1_train/c000_el")
dim(train.data0)
names(train.data0)  = c("peptide", "binder", "HLA")
head(train.data0)

alleles = fread("../../data/SA_data/alleles_SAonly.txt")
dim(alleles)
allele = alleles$cell_line
  
table(allele %in% train.data0$HLA)

t = train.data0[which(train.data0$HLA %in% allele),]
dim(t)
head(t)
table(t$HLA)
table(t$binder, t$HLA) #total peptides per HLA
```

## Duplicated Peptide tables
```{r}

# Number of peptides which are part of duplication
sum(duplicated(t$peptide) | duplicated(t$peptide, fromLast = TRUE))
t1 = t[which(duplicated(t$peptide) | duplicated(t$peptide, fromLast = TRUE)),]
t1[1:15,]
table(t1$binder, t1$HLA)
table(t1$HLA)

round(table(t1$binder, t1$HLA)[2,]/table(t1$HLA),3) #Proportion of duplicated peptides that are binders
#table(t[which(t$binder==1),]$HLA)


# Total number of unique peptides 
length(unique(t1$peptide))

# How many HLA does a unique peptide bind to? (i.e. 752 peptides are duplicated twice/bind to 2 HLA)
table(table(t1$peptide))

# How many peptides does a pair of HLA share? (i.e.273 grouping of HLA share only 1 peptide)
t2 = ddply(t1, .(peptide), summarise, HLApair=paste0(HLA, collapse=","))
#t2
head(t2)
dim(t2)
table(t2$HLApair)
table(table(t2$HLApair))
table1 = table(t2$HLApair) 
table1[table1>5]#checking that table was made correctly (should be 6 peptides)
t2[which(t2$HLApair =="HLA-A01:01,HLA-A02:01"),] 

df1 = data.frame(hlapair = names(table1), num_pep=as.numeric(table1))
df1$hla_in_group = str_count(df1$hlapair, ",")+1
df1$hlapair[which(df1$num_pep < 5)] = ""
ggplot(df1, aes(x=hla_in_group, y=num_pep)) + geom_point() + 
  geom_text_repel(aes(label = hlapair), size = 3.5, 
                     min.segment.length = 0, 
                     box.padding = unit(0.25, "lines"),
                     point.padding = unit(0.25, "lines")) +
  ggtitle(" ")

```

## Hierarchical Clustering
```{r}
head(t)
dim(t)
hla.list = unique(t$HLA)
#hla.list = hla.list[1:5]

dist.ds = matrix(NA, nrow = length(hla.list), ncol = length(hla.list))

for(i in 1:length(hla.list)){
  #print(paste("i=", i, sep  =""))
  #print(hla.list[i])
  #k = i+1
  #if(k<=length(hla.list)){
  #  for(j in k:length(hla.list)){
  for(j in 1:length(hla.list)){
    #print(j)
    #print(hla.list[j])
    
    hla1 = t[which(t$HLA==hla.list[i]),]
    dim(hla1)
    hla2 = t[which(t$HLA == hla.list[j]),]
    dim(hla2)
    
    over = merge(hla1, hla2, by = c("peptide", "binder"))
    unique(over$peptide)
    
    # m = number of of peptide to HLA1 (including overlap)
    m = nrow(hla1)
    # n = # peptide not bind to HLA1 (total - m)
    n = nrow(t) - nrow(hla1)
    # k = # peptide to HLA2 (including overlap)
    k = nrow(hla2)
    # q = # peptide bind to HLA1 and HLA2 (overlap)
    q = nrow(over)-1
    
    #phyper(q,m,n,k, lower.tail = TRUE, log.p = FALSE)
    lowtail = phyper(q,m,n,k, lower.tail = TRUE, log.p = FALSE)
    
    dist.ds[i,j] = lowtail
    }
  #}else{
    #print("reached end of row")
    #next 
  #}
  
}

head(dist.ds)
rownames(dist.ds) = colnames(dist.ds) = hla.list

#change lower tail to p-value
dist.ds.pval = 1- dist.ds

#truncate p-vaue 
dist.ds.tilde = ifelse(dist.ds.pval < 1E-10, 1E-10, dist.ds.pval)
sum(dist.ds.tilde<=1E-10)

#take 10+log10 transofmration
dist.ds.neglog10  = -log10(dist.ds.tilde)
dist.ds.log10  = log10(dist.ds.tilde)
dist.ds.tilde1 = 10+dist.ds.log10
head(dist.ds.tilde1)
sum(is.na(dist.ds.tilde1))

dist.ds1 = as.dist(dist.ds.tilde1, upper =FALSE)
head(dist.ds1)



# Heirarchcal clustering
# Dissimilarity matrix
#d <- dist(dist.ds, method = "euclidean")
# Hierarchical clustering using Complete Linkage
hc1 <- hclust(dist.ds1, method = "complete" )
plot(hc1)
library(ggdendro)
ggdendrogram(hc1, rotate = TRUE, size = 10, theme_dendro = TRUE)

# install.packages("dendextend")
# install.packages("circlize")
library(dendextend)
library(circlize)
hc2 = as.dendrogram(hc1)
hc3 <- hc2 %>%
  color_branches(k = 128) %>%
  color_labels(k = 128)
circlize_dendrogram(hc3,
                    labels_track_height = NA,
                    dend_track_height = 0.3)

#export to pdf
pdf("../../../figures/SA/HLA_clustering.pdf")
circlize_dendrogram(hc3,
                    labels_track_height = NA,
                    dend_track_height = 0.3)
dev.off()

hla1 = t[which(t$HLA=="HLA-B44:02"),]
dim(hla1)
hla2 = t[which(t$HLA == "HLA-B44:28"),]
dim(hla2)
    
over = merge(hla1, hla2, by = c("peptide", "binder"))
unique(over$peptide)

```



## Looking at the group of HLA with lots of peptides 
```{r}
group = names(table1[table1>5])
for(i in length(group)){
  g1 = group[[i]]
  print(g1)
  t3 = t2[which(t2$HLApair ==g1),] 
  print(paste("number of peptides  = ", nrow(t3), 
              sep  =""))
  
  t3$pep_length = nchar(t3$peptide)
  print(table(t3$pep_length))
}

```


